

# DISCLAMER:
These are recent notes that are not intended to be comprehensive.

```{r, message=FALSE}
library(geneorama)
```

## CLIPBOARD ACCESS FOR LINUX... FINALLY
```{r, eval=FALSE}

## Install xclip first:
"$ sudo apt-get install xclip"

# Install from CRAN
install.packages("clipr")

# Or try the development version
devtools::install_github("mdlincoln/clipr")

# library("clipr")
cb <- clipr::read_clip()

cb <- write_clip(c("Text", "for", "clipboard"))
cb <- write_clip(c("Text", "for", "clipboard"), breaks = ", ")

## Future use in geneorama?
## Nice fread example
con <- pipe("xclip -o -selection clipboard")
content <- scan(con, what = character(), sep = "\n", blank.lines.skip = FALSE, quiet = TRUE)
fread(paste(content, collapse = "\n"))
close(con)
```

## Record your screen in Linux
This isn't R, but it's amazing.  Use this code to record your screen in Linux.  Source: http://www.commandlinefu.com/commands/browse

```{r, eval=FALSE}
"ffmpeg -f x11grab -r 25 -s 800x600 -i :0.0 /tmp/outputFile.mpg"
```

## Add alpha to a plot & Smooth Scatter example
```{r}

# source:
# http://menugget.blogspot.com/2012/04/adding-transparent-image-layer-to-plot.html

add_alpha <- function(COLORS, ALPHA){
    if(missing(ALPHA)) stop("provide a value for alpha between 0 and 1")
    RGB <- col2rgb(COLORS, alpha=TRUE)
    RGB[4,] <- round(RGB[4,]*ALPHA)
    NEW.COLORS <- rgb(RGB[1,], RGB[2,], RGB[3,], RGB[4,], maxColorValue = 255)
    return(NEW.COLORS)
}
cols <- c('transparent','blue','yellow','red','darkred')
colramp <- colorRampPalette(add_alpha(cols, .5), alpha=T)
df <- data.table(x=rnorm(100), y=rnorm(100))
df[ , plot(x,y)]
df[ , smoothScatter(x,y,colramp = colramp, add=TRUE,
                        nbin = c(300, 300), bandwidth = c(.2, .2),
                        transformation=function(x) sqrt(x))]
```

## Get google map and plot it

```{r}
library(ggmap)

set_project_dir("geneorama")
infile <- "doc/ggmap_chicago.Rds"

## LOAD DATA
## Also, clean up variable names, and convert dates
if(!file.exists(infile)){
	mapdata <- get_map("Chicago, Illinois", zoom=10)
	saveRDS(mapdata, infile)
}
mapdata <- readRDS(infile)

ggmap(mapdata)
str(mapdata)
# saveRDS(mapdata, "ggmap_data.Rds")
```

## Make a color palette

```{r}
library()
pal <- leaflet::colorQuantile("Greens", NULL, n = 6)
pal <- leaflet::colorNumeric('PuBuGn', 10)

df <- data.table(x=rnorm(1000), y=rnorm(1000))
vals <- df[,1/(3+(x+y)^2)]
pal <- leaflet::colorNumeric('PuBuGn', range(vals))
df[ , plot(y~x, pch=19, col=pal(vals), cex=5)]
```

## Display color palettes (color blind friendly)
```{r}
RColorBrewer::display.brewer.all(colorblindFriendly=TRUE)
RColorBrewer::brewer.pal.info

```

## LEAFLET EXAMPLE FROM GIS STACK EXCHANGE

http://gis.stackexchange.com/questions/168886/r-how-to-build-heatmap-with-the-leaflet-package/203623#203623

Modified to use RCurl and adding elements from example from food-inspections-model (recent branch)

```{r}
set_project_dir("geneorama")

## INITIALIZE
loadinstall_libraries(c("leaflet", "data.table", "sp", "rgdal", "KernSmooth", "RCurl"))
# library("maptools")

inurl <- "https://data.cityofchicago.org/api/views/22s8-eq8h/rows.csv?accessType=DOWNLOAD"
infile <- "doc/mvthefts.Rds"

## LOAD DATA
## Also, clean up variable names, and convert dates
if(!file.exists(infile)){
    # download.file(url = inurl, destfile = infile)
	dat <- fread(RCurl::httpGET(inurl)[1])
	setnames(dat, tolower(colnames(dat)))
	setnames(dat, gsub(" ", "_", colnames(dat)))
	dat <- dat[!is.na(longitude)]
	dat[ , date := as.IDate(date, "%m/%d/%Y")]
	saveRDS(dat, infile)
}
dat <- readRDS(infile)

## MAKE CONTOUR LINES
## Note, bandwidth choice is based on MASS::bandwidth.nrd()
kde <- bkde2D(dat[ , list(longitude, latitude)],
              bandwidth=c(.0045, .0068), gridsize = c(100,100))
CL <- contourLines(kde$x1 , kde$x2 , kde$fhat)

## EXTRACT CONTOUR LINE LEVELS
LEVS <- as.factor(sapply(CL, `[[`, "level"))
NLEV <- length(levels(LEVS))

## CONVERT CONTOUR LINES TO POLYGONS
pgons <- lapply(1:length(CL), function(i)
    Polygons(list(Polygon(cbind(CL[[i]]$x, CL[[i]]$y))), ID=i))
spgons = SpatialPolygons(pgons)

## MAPBOX INFO
MAPBOX_STYLE_TEMPLATE <- paste0("https://api.mapbox.com/styles/v1/coc375492/",
                                "cirqd7mgf001ygcnombg4jtb4/tiles/256/{z}/{x}/{y}",
                                "?access_token=pk.eyJ1IjoiY29jMzc1NDkyIiwiYSI6ImN",
                                "pcnBldzVqMTBmc3J0N25rZTIxZ3ludDIifQ.DgJIcLDjC1h9MtT8CaJ-pQ")
mb_attribution <- paste("© <a href='https://www.mapbox.com/about/maps/'>Mapbox</a> ",
                        "© <a href='http://www.openstreetmap.org/about/'>OpenStreetMap</a>")


## Leaflet map with points and polygons
## Note, this shows some problems with the KDE, in my opinion...
## For example there seems to be a hot spot at the intersection of Mayfield and
## Fillmore, but it's not getting picked up.  Maybe a smaller bw is a good idea?

dat[ , LABEL := paste0(date, " | ", location_description, " | arrest:", arrest)]

```
```{r}
# pal <- leaflet::colorQuantile("Greens", NULL, n = NLEV)
pal <- leaflet::colorFactor("Greens", NULL, levels = NLEV)
pal <- leaflet::colorFactor("Greens", levels = -NLEV:NLEV)
# pal <- leaflet::colorNumeric('PuBuGn', -5:NLEV)
# pal <- leaflet::colorNumeric('PuOr', NLEV:-1)
leaflet(spgons) %>% 
	# addProviderTiles("CartoDB.Positron") %>%
	addTiles(urlTemplate = MAPBOX_STYLE_TEMPLATE, attribution = mb_attribution) 	%>%
	# addPolygons(color = heat.colors(NLEV, NULL)[LEVS], weight=1, fillOpacity=.25) %>%
	addPolygons(color = pal(as.numeric(LEVS)), weight=1, fillOpacity=.25) %>%
    addCircles(lng = ~longitude, lat = ~latitude, weight = 3, popup = ~LABEL,
    		   data = dat, radius = .5, opacity = .4, 
    		   col = ifelse(dat$arrest=="true", "yellow", "red")) %>% 
	addLegend(pal = pal, 
              values = LEVS, 
              position = "bottomright", 
              title = "Crime Intensity") %>% 
	addLegend(colors = c("yellow", "red"),
              labels = c("true", "false"), 
              position = "bottomleft", 
              title = "Arrest")

```




```{r}
## Uncomment to save results
# library(maptools)
# spdf <- SpatialPolygonsDataFrame(spgons, as.data.frame(LEVS), match.ID = F)
# dircreate("mapdata")
# writePolyShape(spdf, "mapdata/any_name")
```



## Stat density

```{r}
geneorama::loadinstall_libraries(c("geneorama", "ggmap", "ggplot2"))

set_project_dir("geneorama")
infile <- "doc/ggmap_chicago.Rds"

## LOAD MAP DATA
## Also, clean up variable names, and convert dates
if(!file.exists(infile)){
	mapdata <- get_map("Chicago, Illinois", zoom=10)
	saveRDS(mapdata, infile)
}
mapdata <- readRDS(infile)

## LOAD CRIME DATA
inurl <- "https://data.cityofchicago.org/api/views/22s8-eq8h/rows.csv?accessType=DOWNLOAD"
infile <- "doc/mvthefts.Rds"

## LOAD DATA
## Also, clean up variable names, and convert dates
if(!file.exists(infile)){
    # download.file(url = inurl, destfile = infile)
	dat <- fread(RCurl::httpGET(inurl)[1])
	setnames(dat, tolower(colnames(dat)))
	setnames(dat, gsub(" ", "_", colnames(dat)))
	dat <- dat[!is.na(longitude)]
	dat[ , date := as.IDate(date, "%m/%d/%Y")]
	saveRDS(dat, infile)
}
dat <- readRDS(infile)

## Crime points
cpts <- dat[,list(lon=longitude, lat=latitude)]

## Density plot (built into ggplot)
base_plot <- ggmap(mapdata)
base_plot + stat_density2d(data=cpts, aes(group=1), color = 4)

## Contour lines - Build kernel density
rng <- unname(unlist(cpts[ , list(range(lon), range(lat))]))
cdens <- MASS::kde2d(x = cpts$lon, y = cpts$lat, h = .03, n = 100, lims = rng)
cdens_dt <- data.table(z=melt(cdens$z))
setnames(cdens_dt, c("x", "y", "z"))
cdens_dt[ , x:=cdens$x[cdens_dt$x]]
cdens_dt[ , y:=cdens$y[cdens_dt$y]]
cdens_dt

## Contour lines - Plot without base layer (proof of concept)
ggplot(data = cdens_dt, aes(x,y,z=z)) + stat_contour(aes(x,y,z=z))
## Contour lines
base_plot + stat_contour(data = cdens_dt, aes(x,y,z=z, group=1))

base_plot + stat_contour(data = cdens_dt, aes(x,y,z=z, group=1)) +
    annotate("text", x = -87.825, y = 41.73, label=paste0("Chicago"), size=8)

base_plot + stat_contour(data = cdens_dt, 
                         aes(x,y,z=z, group=1, colour = ..level..), size=1)

base_plot + 
    stat_contour(data = cdens_dt, geom="polygon", alpha=.2,
                 aes(x,y,z=z, group=1, fill = ..level..)) +
    annotate("text", x = -87.825, y = 41.73,
             label=paste0("Burglary density\nin 2013"), size=7)
```

